---
title: "Tutorial1"
author: "Danny Muziga"
date: "March 30, 2016"
output: html_document
---

This is my first Markdown document 

let's load some data 

# Question 1 : Merging the training and the test sets to create one data set 

```{r}

# Load activity labels + features
activityNames <- read.table("activity_labels.txt", header = FALSE)
## CHANGING THE ACtivityName from the  factor to character
activityNames$V2 <- as.character(activityNames$V2)

featureNames <- read.table("features.txt")


# Reading the data set for the testdatain R
testset <- read.table("test/X_test.txt", header = FALSE)

dim (testset)
testsetname <- read.table(file = "test/y_test.txt", 
                          header = FALSE)
                          
dim(testsetname)

subject_test <- read.table(file = "test/subject_test.txt", 
                           header = FALSE)
 dim(subject_test)                    
# Merging the data 

testdataset <- cbind(testset, testsetname,subject_test)

dim(testdataset)

#Reading the data set for the training datasetin R
trainset <- read.table( 
        file = "train/X_train.txt",
        header = FALSE)

dim(trainset)
trainsetname <- read.table(file = "train/y_train.txt", 
                           header = FALSE)
                           

dim(trainsetname)

subject_train<- read.table(file = "train/subject_train.txt", 
                           header = FALSE)
   
dim(subject_train)


#combine training and test dataset
subject <- rbind(subject_train, subject_test)
activity <- rbind(trainsetname,testsetname)
features <- rbind(trainset,testset )

#Naming Column 
colnames(features) <- featureNames[,2]
colnames(activity) <- "Activity"
colnames(subject) <- "Subject"


#Merge Data
mergeData <- cbind(features,activity, subject)

dim(mergeData)

```
# Question 2 : Extracts only the measurements on the mean and standard deviation for each measurement. 
```{r}
# return the position of the value that have mean and  std in the feature column 
featuresNeeded <- grep(".*mean.*|.*std.*", names(mergeData), ignore.case = TRUE)

#returning the column name for mean and std (standard deviation) including the Activity (column number = 562) and Subject (column number = 563)
meansSTD <- c(featuresNeeded, 562,563)

# extract the the column that have the mean and STD
extractedFeatures <-mergeData[, meansSTD ]

class(featuresNeeded)

dim(extractedFeatures)

```
# Question 3: Uses Descriptive activity names to name the activities in the data set 

```{r}

# Uses descriptive Activity names to name the activities in the dataset

colnames(extractedFeatures) = gsub('-mean', 'Mean',colnames(extractedFeatures))
colnames(extractedFeatures) = gsub('-std', 'Standard_Deviation',colnames(extractedFeatures))
colnames(extractedFeatures) <- gsub('[-()]', '', colnames(extractedFeatures))

```

# Question 4 : Appropriately labels the data set with descritive variable Names
```{r}

#Appropriately labels the data set with descriptive variable names 
# we remove the Abreviation to put on full name.

colnames(extractedFeatures) <- gsub('Acc', 'Accelerometer',colnames(extractedFeatures))
colnames(extractedFeatures) <- gsub('Gyro', 'Gyroscope',colnames(extractedFeatures))
colnames(extractedFeatures) <- gsub('BodyBody', 'Body',colnames(extractedFeatures))
colnames(extractedFeatures) <- gsub('Subject1', 'Subject',colnames(extractedFeatures))

```
# Question 5:  From the data set in Question 4, create a second independent tidy data set with the average of each variable for each activity and each subject 

```{r}

library(reshape2)

#I firstly make the Activity and Subject as our Id then we  use dcast to calculate the mean 
newtidy <- melt(extractedFeatures, id.vars = c("Activity", "Subject"))

newtidy <- dcast(newtidy, Activity + Subject ~ variable, fun.aggregate = mean)
newtidy <- newtidy[order(newtidy$Subject,newtidy$Activity),]

#write the result to a file 
write.table(newtidy, file = "TidyDataSet.csv", sep = ",",col.names = TRUE)

```
